Real-Time Activity Classification by Matched Filtering using Body-Worn Accelerometers

### 2016-09-26 notes/paper/paper.tex
Incremetnal changes.

To compute the matched-filter output for two dimensions, we generalize \eqref{eq:matched_filter} by augmenting our reference and signal with both dimensions appropriately for normalization and adding together the cross correlations as:
%
\begin{equation} \label{eq:cross_correlation_2}
\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k := \frac{(\textbf{x}^1 \star \textbf{y}^1)_k + (\textbf{x}^2 \star \textbf{y}^2)_k}{||[ \textbf{x}^1, \textbf{x}^2 ]|| \ || [ \widetilde{\textbf{y}}_k^1, \widetilde{\textbf{y}}_k^2 ] || },
\end{equation}
%
for $ k = 0,1,...,N-M-1 $. \\
%
\begin{equation} \label{eq:matched_filter_2}
MF_{\textbf{x}}^2(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k \right \}
\end{equation}
%
This works because both dimensions are always in phase with each other for any given activity and therefore the match will happen in sync. Here, the superscripts $1$ and $2$ represent the primary and secondary principal components of our data respectively.
%
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of the signal by filtering to mitigate any possible distortion effects due to aliasing followed by downsampling.
Because real-time performance is important, we decimate our signal for the purpose of improving our throughput.

@@

\begin{figure}[!ht]
   \centering
   \includegraphics[width=0.45\textwidth]{32Hz_4sec_8downsample_2Reference_Pulse.PNG}
   \caption{Oscilloscope output used for measuring the processing time of the algorithms on the MSP432. This is a case setup of 32Hz sampling frequency over a 4 second time window with 2 references at decimation 8 showing the voltage output over time. Execution time is measured from the start to end of the ``high'' (3.3 Volts) time interval.}
   \label{fig:time_measure}
\end{figure}
%
The ``Analog Discovery 100MSPS USB Oscilloscope and Logic Analyzer'' oscilloscope was used for measuring the timing of the algorithms on the microcontroller.
When testing the execution speed, one of the General-Purpose Input/Output (GPIO) pins was connected to a Light-emitted diode (LED) that activates at 3.3 Volts (set high).
This is activated immediately before the beginning of decimation (start of algorithm processing) and deactivated immediately at the end of the matched-filtering (end of algorithm processing).
The execution time is then recorded on the oscilloscope to determine how long the GPIO was set high.
An example can be seen in Figure \ref{fig:time_measure}.

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds with timings measured at sampling frequencies of $16, 32, 62, 64, 75, 100,$ and $128$ Hz.}
   \label{fig:timing}
\end{figure}
%
Because the accelerometer was not connected to the microcontroller at the time of writing this paper, randomized data was used for the purpose of analyzing the throughput performance on the device while varying the parameters which accurately models the processing load we would expect when we do have the accelerometer connected to the microcontroller.
Figure \ref{fig:multiple_comparison} shows the execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process and operating at decimation $2$.
Because the FFT dominates the processing load, and was optimized for even-powered factors, one expects best throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even-valued, non-radix 2, time-lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than that the $50$Hz.
Figure \ref{fig:timing_refs} shows the execution time over a $4.0$ second time window at decimation $8$ while varying the sampling frequency and number of references. Because the pre-processing execution time does not change with the number of references, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\% when processing for a single reference.
%
\section{Conclusion}
We have demonstrated excellent activity classification capability of the matched-filter method using data obtained from a Kinetisense tri-axis accelerometer sensor.
We have shown that real-time processing can be achieved with timing results obtained from the MSP432 microcontroller.
Timing results have demonstrated that radix 2 time-window lengths and frequency values result in faster processing times as expected given that the FFT dominates the processing time.
We increased our throughput with the use of data-compression techniques, namely decimation and dimensionality reduction, which reduced our computational requirements.
The references used in the matched-filter are obtained from the training module where a training set is provided by the user through some specified activity and the best representative motion signature of the individual is extracted through the use of an instance-based learning algorithm.

>>>>

Classification methods that are achieved from multi-dimensional feature spaces have been explored \cite{survey_multi_view_machine_learning}.
In this paper, we compute the matched-filter output for two dimensions by generalizing \eqref{eq:matched_filter} by augmenting our reference and signal with both dimensions appropriately for normalization and adding together the cross correlations as:
%
\begin{equation} \label{eq:cross_correlation_2}
\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k := \frac{(\textbf{x}^1 \star \textbf{y}^1)_k + (\textbf{x}^2 \star \textbf{y}^2)_k}{||[ \textbf{x}^1, \textbf{x}^2 ]|| \ || [ \widetilde{\textbf{y}}_k^1, \widetilde{\textbf{y}}_k^2 ] || },
\end{equation}
%
for $ k = 0,1,...,N-M-1 $. \\
%
\begin{equation} \label{eq:matched_filter_2}
MF_{\textbf{x}}^2(\textbf{y}) := \underset{k \in [0, M-N)}{max} \left \{\widehat{(\textbf{x}^{1,2} \star \textbf{y}^{1,2})}_k \right \}
\end{equation}
%
This works because both dimensions are always in phase with each other for any given activity and therefore the match will happen in sync. Here, the superscripts $1$ and $2$ represent the primary and secondary principal components of our data respectively.
%
\subsection{Decimation}
Decimation is the first step applied to our data before any additional processing takes place.
Decimation is a process of reducing the sampling rate of the signal by filtering to mitigate any possible distortion effects due to aliasing followed by downsampling.
Because real-time performance is important, we decimate our signal for the purpose of improving our throughput.

@@

\begin{figure}[!ht]
   \centering
   \includegraphics[width=0.45\textwidth]{reference_pulse_32Hz_4sec_8downsample_2.png}
   \caption{Oscilloscope output used for measuring the processing time of the algorithms on the MSP432. This is a case setup of 32Hz sampling frequency over a 4 second time window with 2 references at decimation 8 showing the voltage output over time. Execution time is measured from the start to end of the ``high'' (3.3 Volts) time interval.}
   \label{fig:time_measure}
\end{figure}
%
The ``Analog Discovery 100MSPS USB Oscilloscope and Logic Analyzer'' oscilloscope was used for measuring the timing of the algorithms on the microcontroller.
When testing the execution speed, one of the General-Purpose Input/Output (GPIO) pins was connected to a Light-emitted diode (LED) that activates at 3.3 Volts (set high).
This is activated immediately before the beginning of decimation (start of algorithm processing) and deactivated immediately at the end of the matched-filtering (end of algorithm processing).
The execution time is then recorded on the oscilloscope to determine how long the GPIO was set high.
An example can be seen in Figure \ref{fig:time_measure}.

\begin{figure}[!ht]
   \centering
   \subfloat[\label{fig:multiple_comparison}]{\includegraphics[width=0.45\textwidth]{more_time_plots.png}}
   \quad
   \subfloat[\label{fig:timing_refs}]{\includegraphics[width=0.45\textwidth]{one_and_two_references.png}}
   \caption{Timing results with varying sampling frequencies, time-window lengths, and number of references used for the matched-filter on the MSP432 microcontroller processor. The time window used for (b) is 4.0 seconds with timings measured at sampling frequencies of $16, 32, 62, 64, 75, 100,$ and $128$ Hz.}
   \label{fig:timing}
\end{figure}
%
Because the accelerometer was not connected to the microcontroller at the time of writing this paper, randomized data was used for the purpose of analyzing the throughput performance on the device while varying the parameters which accurately models the processing load we would expect when we do have the accelerometer connected to the microcontroller.
Figure \ref{fig:multiple_comparison} shows the execution time with varying time-window lengths and sampling frequencies while maintaining two references used in the matched-filter process and operating at decimation $2$.
Because the FFT dominates the processing load, and a custom FFT was written specifically for this unit and optimized for even-powered factors, one expects best throughput at radix 2 valued time-window lengths and sampling frequencies followed by other even-valued, non-radix 2, time-lengths and sampling frequencies. Both $62$Hz and $50$Hz have only one factor of $2$ hence the expectation that a sampling frequency of $62$Hz processing time is longer than that of the $50$Hz.
Figure \ref{fig:timing_refs} shows the execution time over a $4.0$ second time window at decimation $8$ while varying the sampling frequency and number of references. Because the pre-processing execution time does not change with the number of references, it's observed that matched-filtering accounts for roughly $80$\% of the processing while pre-processing accounts for the other $20$\% when processing for a single reference.
%
\begin{figure}[!ht]
   \centering
   \includegraphics[width=0.6\textwidth]{flops_graph.png}
   \caption{An estimate of the flops used to accomplish the main processing tasks was made. This was accomplished by tallying all the low level operations with in addition to using the $2.5 N log_2 N$ mflops estimate for real-FFTs from FFTW \cite{FFTW}. Although the FFT used on the MSP432 was custom, off-line implementations of the algorithms made use of FFTW}
   \label{fig:flops_count}
\end{figure}
%
\section{Conclusion}
We have demonstrated excellent activity classification capability of the matched-filter method using data obtained from a Kinetisense tri-axis accelerometer sensor.
We have shown that real-time processing can be achieved with timing results obtained from the MSP432 microcontroller.
Timing results have demonstrated that radix 2 time-window lengths and frequency values result in faster processing times as expected given that the FFT dominates the processing time.
We increased our throughput with the use of data-compression techniques, namely decimation and dimensionality reduction, which reduced our computational requirements.
The references used in the matched-filter are obtained from the training module where a training set is provided by the user through some specified activity and the best representative motion signature of the individual is extracted through the use of an instance-based learning algorithm.
